FlowFPX: Nimble Tools for Debugging Floating-Point Exceptions

### 2023-08-17 paper.tex
Last edits to paper (this round)

\begin{verbatim}
-([-Inf, -Inf])   FT/TrackedFloat.jl:106
momentum_u!       ShallowWaters/rhs.jl:246
rhs_nonlinear!    ShallowWaters/rhs.jl:50
rhs!              ShallowWaters/rhs.jl:14 [inlined]
time_integration  ShallowWaters/time_integration.jl:77
run_model         ShallowWaters/run_model.jl:37
top-level
\end{verbatim}

@@

\begin{verbatim}
^([-1.515f31, 2])          FT/TrackedFloat.jl:138
literal_pow()              intfuncs.jl:325
…
materialize(^)             broadcast.jl:860
top-level getproperty(…)  examples/sw_nan_tf.jl:14
…
\end{verbatim}

@@

We can see in \cref{fig:sw_nan_cstg} that a thick line goes from the top-level call down to a call to \texttt{+} goes down the right side of the graph.
This indicates that the majority of the flows took this path.
That means we might be able to fix the majority of our problems by inspecting that particular flow.
\Cref{fig:sw_inf_cstg} shows us that all but two flows started with a call to \texttt{^}---a promising place to start looking for improvements.

@@

The green lines indicate flows that were \emph{new} in the latter 90\% of the logs.
The red lines indicate flows that were \emph{only in} the first 10\% of the logs.
What this means depends on the particular program; a domain expert can use these clues to e.g. find where instability started in the first part of the program, and where that instability got reinforced later on.
In this case, we can see that \texttt{momentum_u!} only showed up in the beginning of the logs---we might look there for ways to prevent instability from beginning.

@@

\begin{verbatim}
<([NaN, 3.0e6])  at FT/TrackedFloat.jl:193
solve!           at OrdinaryDiffEq/solve.jl:515
…
\end{verbatim}

@@

Fuzzing with \FT{} revealed two places where Finch needed
protection against user input.
The first was when reading an input mesh.\footnote{\urlaccess{https://github.com/paralab/Finch/issues/16}{2023-06-06}}
A \NaN{} injected in the mesh led to a crash further on:
\begin{verbatim}
  BoundsError: attempt to access 1-element
   Vector{Int64} at index [2]
\end{verbatim}
Defensive programming solves the issue.
The second place was in setting bounds for the
solver.\footnote{\urlaccess{https://github.com/paralab/Finch/issues/17}{2023-06-06}}
Here, a \NaN{} could leave bounds uninitialized, leading to a bounds error.
Additionally, \FT{} and \CSTG{}s have been useful for identifying \NaN{}s that
appear in unstable systems.
%% advection2d, high cfl and high T (time) led to NaN during simulation

@@

\FT{} has proven to be useful for more than its creators---we discovered an open issue in the \code{RxInfer.jl} library with an issue revolving around \NaN{} detection.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
Even though we were unable to provide hands-on guidance as the code in question was proprietary, the authors of \code{RxInfer.jl} were able use \FT{} themselves, and within a day they tracked down the location of a \NaN{} gen.
The \code{RxInfer.jl} authors provided us with valuable feedback on \FT{}'s API.

@@

Toronto and McCarthy\cite{torontoPracticallyAccurateFloatingPoint2014} developed a methodology for assessing where arithmetic operations are liable to magnify \fp{} error.
Their work includes error visualization and work on how to avoid error-magnifying ``badlands''.
They focus on rewriting arithmetic expressions to avoid error which could lead to exceptional values, rather than on exceptional values themselves.

@@

The Julia library \texttt{Sherlogs}\footnote{\url{https://github.com/milankl/Sherlogs.jl}} by Milan Klöwer inspired the use of a custom number type to intercept operations on a number.
In contrast to \FT{} which monitors for exceptional values and logs stack traces at interesting points in their lifetime, \texttt{Sherlogs} tracks and reports the range of values seen over the course of a computation.
This is intended to provide insight into whether or not a library could tolerate a lower-precision \fp{} format.

@@

FloatTracker is capable of monitoring all sorts of events in a \fp{} value's lifetime---not just a value going to \NaN{} or \Inf{}.
We spent most of the time engineering and testing \NaN{} tracking and injection.
Adding \Inf{} tracking took a few hours.
Adding other values of interest presents no fundamental difficulties.

@@

While fuzzing is useful for discovering issues, its success rate
is low because \emph{every} \fp{} operation is a candidate
for injection.
Even operations that are already well-defended against \NaN{}s are candidates.
\FT{} could use two sorts of tools for improving injection.
First, fine-grained control to let users decide where not to inject.
Second, tools for understanding the context of an injection point
after the fact.
Program slicing is especially relevant to the latter point (and effectively
what we did by hand when fuzzing Finch~\cref{s:finch}).
For each operation, an expert needs to study the values that feed into it to
decide whether they are protected or not.

>>>>

\begin{verbatim}
-([-Inf, -Inf])   FT/TrackedFloat.jl:106
momentum_u!       SW/rhs.jl:246
rhs_nonlinear!    SW/rhs.jl:50
rhs!              SW/rhs.jl:14 [inlined]
time_integration  SW/time_integration.jl:77
run_model         SW/run_model.jl:37
top-level
\end{verbatim}

@@

\begin{verbatim}
^([-1.515f31, 2])          FT/TrackedFloat.jl:138
literal_pow()              intfuncs.jl:325
…
materialize(^)             broadcast.jl:860
top-level getproperty(...) examples/sw_nan_tf.jl:14
…
\end{verbatim}

@@

We can see in \cref{fig:sw_nan_cstg} that a thick line goes from the top-level call down to a call to \texttt{+} goes down the right side of the graph.
This indicates that the majority of the flows took this path.
That means we might be able to fix the majority of our problems by inspecting that particular flow.
\Cref{fig:sw_inf_cstg} shows us that all but two flows started with a call to \texttt{\^}---a promising place to start looking for improvements.

@@

The green lines indicate flows that were \emph{new} in the latter 90\% of the logs.
The red lines indicate flows that were \emph{only in} the first 10\% of the logs.
What this means depends on the particular program; a domain expert can use these clues to e.g. find where instability started in the first part of the program, and where that instability got reinforced later on.
In this case, we can see that \texttt{momentum\_u!} only showed up in the beginning of the logs---we might look there for ways to prevent instability from beginning.

@@

\begin{verbatim}
<([NaN, 3.0e6])  at FT/TrackedFloat.jl:193
solve!           at ODE/solve.jl:515
…
\end{verbatim}

@@

Fuzzing with \FT{} revealed two places where Finch needed
protection against user input.
The first was when reading an input mesh.\footnote{\urlaccess{https://github.com/paralab/Finch/issues/16}{2023-06-06}}
A \NaN{} injected in the mesh led to a crash further on:

\begin{verbatim}
  BoundsError: attempt to access 1-element
   Vector{Int64} at index [2]
\end{verbatim}

Defensive programming solves the issue.
The second place was in setting bounds for the
solver.\footnote{\urlaccess{https://github.com/paralab/Finch/issues/17}{2023-06-06}}
Here, a \NaN{} could leave bounds uninitialized, leading to a bounds error.
Additionally, \FT{} and \CSTG{}s have been useful for identifying \NaN{}s that
appear in unstable systems.
%% advection2d, high cfl and high T (time) led to NaN during simulation

@@

\FT{} has proven to be useful for more than its creators---we discovered an open issue in the \code{RxInfer.jl} library with an issue revolving around \NaN{} detection.\footnote{\url{https://github.com/biaslab/RxInfer.jl/issues/116}}
Even though we were unable to provide hands-on guidance as the code in question was proprietary, the authors of \code{RxInfer.jl} were able use \FT{} themselves, and within a day they tracked down the location of a \NaN{} gen.

@@

Toronto and McCarthy\cite{torontoPracticallyAccurateFloatingPoint2014} develop a methodology for assessing where arithmetic operations are liable to magnify \fp{} error.
They include error visualization and work on how to avoid error-magnifying ``badlands''.
They focus on rewriting arithmetic expressions to avoid error which could lead to exceptional values, rather than on the exceptional values themselves.

@@

The Julia library \texttt{Sherlogs}~\cite{kMilanklSherlogsJl2021} inspired the use of a custom number type to intercept operations on a number.
In contrast to \FT{} which monitors for exceptional values and logs stack traces at interesting points in their lifetime, \texttt{Sherlogs} tracks and reports the range of values seen over the course of a computation.
This is intended to provide insight into whether or not a library could tolerate a lower-precision \fp{} format.

@@

FloatTracker is capable of monitoring all sorts of events in a \fp{} value's lifetime---not just a value going to \NaN{} or \Inf{}.
For example, \FT{} could be made to monitor extreme values: in some circumstances it would be useful to notice when a value started climbing above or falling below a certain threshold.

We spent most of the time engineering and testing \NaN{} tracking and injection.
Adding \Inf{} tracking took a few hours.
Adding other values of interest presents no fundamental difficulties.

@@

While fuzzing is useful for discovering issues, its success rate
is low because \emph{every} \fp{} operation is a candidate
for injection.
Even operations that are already well-defended against \NaN{}s are candidates.
\FT{} could use two sorts of tools for improving injection.
First, fine-grained control to let users decide where not to inject.
Second, tools for understanding the context of an injection point
after the fact.
Program slicing is especially relevant to the latter point and effectively
what we did by hand when fuzzing Finch (\cref{s:finch}).
For each operation, an expert needs to study the values that feed into it to
decide whether they are protected or not.
